\large{\bf Looking forwards and backwards through locally regulated populations }

### 2023-04-13 forwards-and-backwards.tex
second bit of appendix

Furthermore, let $\Pgen^\infty$ be the generator given by
\begin{equation}
    \label{eq: Limit Generator Definition}
\begin{aligned}
\Pgen^{\infty} G_f(\eta)
    &:= G'(\langle f, \eta \rangle)
                   %\times 
\left\{
                   \big\langle
                        \gamma(x, \smooth{\gamma} \eta(x))
                            \mathcal{B}\left(
                            f(\cdot) r(\cdot, \smooth{r} \eta(\cdot))
                            \right)(x)
                    +
                    f(x)
                        F(x, \smooth{F} \eta(x)),
                        \eta(dx)
                    \big\rangle
                   \right\}\\
               &\qquad {}
               + \frac{\alpha}{2} G''(\langle f, \eta \rangle)
                  %\times 
\left\{
                  \big\langle
                    2\gamma\left( x, \eta \right)
                    r\left(x,\eta \right)
                    %+\mu\left(x,\eta \right)
                    f^2(x),
                    \eta (dx)
                    \big\rangle 
                  \right\}.
\end{aligned}    
\end{equation}
A process $(\eta^{\infty}_t)_{t \geq 0}$ 
satisifes the martingale characterization of equations~\ref{eqn:limiting_mgale_problem}
and~\ref{eqn:limiting_mgale_variation}
if and only if it is a solution 
to the $(\Pgen^{\infty}, \eta^\infty_0)$-martingale problem, i.e.,
if for all such test functions 
$$
    M_t :=
    G_f(\eta^{\infty}_t) - G_f(\eta^\infty_0)
    - \int_{0}^{t}\Pgen^{\infty}G_f(\eta^{\infty}_s)ds
$$
is a martingale (with respect to the natural $\sigma$-field). 
\end{lemma}

@@

For the limiting process,
we now require test functions such that $g(x,u) = 1$ for all $u$ larger than some $u_0$;
As before, recall that $\kappa(\lp)(\cdot) = \lim_{u \to \infty} \lp(\cdot \times [0, u)) / u$
is the ``spatial projection operator'',
and define the probability kernel $\Gamma : \measures \to \flpmeasures$
so that for $\eta \in \measures$,
defining $h(x) = \int_0^\infty (g(x, u) - 1) du$,
\begin{align*}
    F_g(\eta)
    &:=
    \int f(\lp) \Gamma(\eta, d\lp) \\
    &=
    \exp\left(
        \big \langle
            \int_0^\infty (g(x, u) - 1) du,
            \eta(dx)
        \big \rangle
    \right) \\
    &=
    e^{ \langle h(x), \eta(dx) \rangle } .
\end{align*}
i.e., $\Gamma(\eta, \cdot)$ is the distribution of a conditionally Poisson process
with intensity a product of $\eta$ and Lebesgue measure.
It again follows from Lemma~\ref{def: MP definition of limit} that for test functions of this form
\comment{(beware the different definition of $f$ to that lemma)}
the generator of $\eta_t$ is
\begin{align}
    \Pgen^\infty F_g(\eta)
    &=
    F_g(\eta)
    \left\langle
        \gamma(x, \eta)
        \DG( h(\cdot) r(\cdot) )(x)
        + F(x, \eta) h(x)
        + \alpha \gamma(x, \eta) r(x, \eta) h^2(x)
        ,
        \eta(dx)
    \right\rangle .
\end{align}
The generator of $\lp_t$ is $A$, defined in equation~\eqref{eqn:limiting_lookdown_generator},

@@

We now integrate each piece against $\Gamma^N$.
First note that by the product form of $f$,
$$
    \int f(\lp) \sum_{(x,u)\in\lp} \frac{h(x,u)}{g(x,u)} \Gamma^N(\eta,d\lp)
    =
    F_g^N(\eta) \bigg\langle
        \frac{1}{N \hat g(x)} \int_0^N h(x,u) du,
        N \eta(dx)
    \bigg\rangle .
$$
Therefore,
\footnotesize
\begin{align*}
&
    \int A^N_1f(\lp) \Gamma^N(\eta,d\lp) 
\\ &\qquad =
    F_g^N(\eta) \bigg\langle \frac{c_\theta(x,\eta)}{\hat g(x)} \int_0^N \Bigg\{
        \int_u^N \Bigg(
            \frac{1}{2} g(x,v_1) \int_{\IR^d}(g(y,u)-g(x,u))q^m_{\theta}(x,dy,\eta)
        \Bigg) dv_1
    \Bigg\} du, \eta(dx) \bigg \rangle
\\ &\qquad =
    F_g^N(\eta) \bigg\langle \frac{c_\theta(x,\eta)}{\hat g(x)} \Bigg\{
    \int_{\IR^d} \left\{
        \int_{0}^{N} \int_u^N g(x,v_1) (g(y,u)-g(x,u)) dv_1 du
    \right\} q^m_{\theta}(x,dy,\eta)
    , \eta(dx) \bigg \rangle .
\end{align*}
\normalsize
For the second generator, we have
\footnotesize
\begin{align*}
&
    \int A^N_2 f(\lp) \Gamma^N(\eta, d\lp)
\\ &\qquad =
    F_g^N(\eta) \bigg\langle \frac{c_\theta(x,\eta)}{\hat g(x)} \int_0^N g(x, u) \Bigg\{
        \int_u^N\Bigg(\int_{\IR^d}
        \left(\frac{g(y,v_1)+g(x,v_1)}{2}-1\right) q^m_{\theta}(x,dy,\eta)
        \Bigg)dv_1 \Bigg\}
    , \eta(dx) \bigg\rangle
\\ &\qquad =
    F_g^N(\eta) \bigg\langle \frac{c_\theta(x,\eta)}{\hat g(x)}
    \int_{\IR^d}\Bigg\{ \int_{0}^{N}
    \int_u^N g(x,u) \left(
        g(y,v_1)+g(x,v_1)-2
    \right) dv_1 du\Bigg\} q^m_{\theta}(x,dy,\eta)
    , \eta(dx) \bigg\rangle .
\end{align*}  
\normalsize
For the third generator we have that, since $g(x,N) = 1$,
\footnotesize
\begin{align*}
&
    \int A^N_3f(\xi)\Gamma^N(\eta,d\lp)
\\ & \qquad =
    F_g^N(\eta) \bigg\langle \frac{1}{\hat g(x)}
    \int_0^N
        \left( c_\theta(x,\eta) u^2 - b_\theta(x,\eta) u \right) \partial_u g(x,u)
    du
    , \eta(dx) \bigg\rangle 
\\ & \qquad =
    F_g^N(\eta) \bigg\langle \frac{1}{\hat g(x)}
    \int_0^N
        \left( b_\theta(x,\eta) - 2 c_\theta(x,\eta) u \right) (g(x,u) - 1)
    du
    , \eta(dx) \bigg\rangle .
\end{align*}
\normalsize

@@

>>>>

Furthermore, let $\Pgen^\infty$ be the generator given by
\begin{equation}
    \label{eq: Limit Generator Definition}
\begin{aligned}
\Pgen^{\infty} G_f(\eta)
    &:= G'(\langle f, \eta \rangle)
                   %\times 
\left\{
                   \big\langle
                        \gamma(x, \eta)
                            \mathcal{B}\left(
                            f(\cdot) r(\cdot, \eta)
                            \right)(x)
                    +
                    f(x) F(x, \eta),
                    \eta(dx)
                    \big\rangle
                   \right\}\\
               &\qquad {}
               + \frac{\alpha}{2} G''(\langle f, \eta \rangle)
                  %\times 
\left\{
                  \big\langle
                    2\gamma\left( x, \eta \right)
                    r\left(x,\eta \right)
                    %+\mu\left(x,\eta \right)
                    f^2(x),
                    \eta (dx)
                    \big\rangle 
                  \right\}.
\end{aligned}    
\end{equation}
A process $(\eta^{\infty}_t)_{t \geq 0}$ 
satisifes the martingale characterization of equations~\ref{eqn:limiting_mgale_problem}
and~\ref{eqn:limiting_mgale_variation}
if and only if it is a solution 
to the $(\Pgen^{\infty}, \eta^\infty_0)$-martingale problem, i.e.,
if for all such test functions 
$$
    M_t :=
    G_f(\eta^{\infty}_t) - G_f(\eta^\infty_0)
    - \int_{0}^{t}\Pgen^{\infty}G_f(\eta^{\infty}_s)ds
$$
is a martingale (with respect to the natural $\sigma$-field). 
\end{lemma}

@@

@@

We now integrate each piece against $\Gamma^N$.
First note that by the product form of $f$,
$$
    \int f(\lp) \sum_{(x,u)\in\lp} \frac{\ell(x,u)}{g(x,u)} \Gamma^N(\eta,d\lp)
    =
    F_g^N(\eta) \bigg\langle
        \frac{1}{N \hat g(x)} \int_0^N \ell(x,u) du,
        N \eta(dx)
    \bigg\rangle .
$$
Therefore,
\footnotesize
\begin{align*}
&
    \int A^N_1f(\lp) \Gamma^N(\eta,d\lp) 
\\ &\qquad =
    F_g^N(\eta) \bigg\langle \frac{c_\theta(x,\eta)}{\hat g(x)} \int_0^N \Bigg\{
        \int_u^N \Bigg(
            \frac{1}{2} g(x,v_1) \int_{\IR^d}(g(y,u)-g(x,u))q^m_{\theta}(x,dy,\eta)
        \Bigg) dv_1
    \Bigg\} du, \eta(dx) \bigg \rangle
\\ &\qquad =
    F_g^N(\eta) \bigg\langle \frac{c_\theta(x,\eta)}{\hat g(x)} \Bigg\{
    \int_{\IR^d} \left\{
        \int_{0}^{N} \int_u^N g(x,v_1) (g(y,u)-g(x,u)) dv_1 du
    \right\} q^m_{\theta}(x,dy,\eta)
    , \eta(dx) \bigg \rangle .
\end{align*}
\normalsize
For the second generator, we have
\footnotesize
\begin{align*}
&
    \int A^N_2 f(\lp) \Gamma^N(\eta, d\lp)
\\ &\qquad =
    F_g^N(\eta) \bigg\langle \frac{c_\theta(x,\eta)}{\hat g(x)} \int_0^N g(x, u) \Bigg\{
        \int_u^N\Bigg(\int_{\IR^d}
        \left(\frac{g(y,v_1)+g(x,v_1)}{2}-1\right) q^m_{\theta}(x,dy,\eta)
        \Bigg)dv_1 \Bigg\}
    , \eta(dx) \bigg\rangle
\\ &\qquad =
    F_g^N(\eta) \bigg\langle \frac{c_\theta(x,\eta)}{\hat g(x)}
    \int_{\IR^d}\Bigg\{ \int_{0}^{N}
    \int_u^N g(x,u) \left(
        g(y,v_1)+g(x,v_1)-2
    \right) dv_1 du\Bigg\} q^m_{\theta}(x,dy,\eta)
    , \eta(dx) \bigg\rangle .
\end{align*}  
\normalsize
For the third generator we have that, since $g(x,N) = 1$,
\footnotesize
\begin{align*}
&
    \int A^N_3f(\xi)\Gamma^N(\eta,d\lp)
\\ & \qquad =
    F_g^N(\eta) \bigg\langle \frac{1}{\hat g(x)}
    \int_0^N
        \left( c_\theta(x,\eta) u^2 - b_\theta(x,\eta) u \right) \partial_u g(x,u)
    du
    , \eta(dx) \bigg\rangle 
\\ & \qquad =
    F_g^N(\eta) \bigg\langle \frac{1}{\hat g(x)}
    \int_0^N
        \left( b_\theta(x,\eta) - 2 c_\theta(x,\eta) u \right) (g(x,u) - 1)
    du
    , \eta(dx) \bigg\rangle .
\end{align*}
\normalsize

@@

For the limiting process,
we now require test functions such that $g(x,u) = 1$ for all $u$ larger than some $u_0$;
As before, recall that $\kappa(\lp)(\cdot) = \lim_{u \to \infty} \lp(\cdot \times [0, u)) / u$
is the ``spatial projection operator'',
and define the probability kernel $\Gamma : \measures \to \flpmeasures$
so that for $\eta \in \measures$,
defining $h(x) = \int_0^\infty (g(x, u) - 1) du$,
\begin{align*}
    F_g(\eta)
    &:=
    \int f(\lp) \Gamma(\eta, d\lp) \\
    &=
    \exp\left(
        \big \langle
            \int_0^\infty (g(x, u) - 1) du,
            \eta(dx)
        \big \rangle
    \right) \\
    &=
    e^{ \langle h(x), \eta(dx) \rangle } .
\end{align*}
i.e., $\Gamma(\eta, \cdot)$ is the distribution of a conditionally Poisson process
with intensity a product of $\eta$ and Lebesgue measure.
It again follows from Lemma~\ref{def: MP definition of limit} that for test functions of this form
\comment{(beware the different definition of $f$ to that lemma)}
the generator of $\eta_t$ is
\begin{align} \label{eqn:pgen_infty}
    \Pgen^\infty F_g(\eta)
    &=
    F_g(\eta)
    \left\langle
        \gamma(x, \eta)
        \DG( h(\cdot) r(\cdot) )(x)
        + F(x, \eta) h(x)
        + \alpha \gamma(x, \eta) r(x, \eta) h^2(x)
        ,
        \eta(dx)
    \right\rangle .
\end{align}
The generator of $\lp_t$ is $A$, defined in equation~\eqref{eqn:limiting_lookdown_generator}.

Before proving Lemma~\ref{lem:averaged_generators_limit},
we recall an important equality for conditionally
Poisson point processes (\cite{kurtz/rodrigues:2011} Lemma A.3).

\begin{lemma} \label{lem:poisson_eqn}
If $\lp = \sum_{i}\delta_{Z_i}$ is a Poisson random measure with mean measure $\nu$, 
then for $\ell \in L^{1}(\nu)$ and $g\geq0$ with $\log g \in L^{1}(\nu)$,
\begin{equation}
\IE\left[\sum_{j} \ell(Z_j) \prod_{i}g(Z_i)\right] = \int \ell g d\nu e^{\int (g-1) d \nu}.
\end{equation}
\end{lemma}

\begin{proof}[of Lemma~\ref{lem:averaged_generators_limit}]
By Lemma~\ref{lem:poisson_eqn},
$$
    \int f(\lp) \sum_{(x,u)\in\lp} \frac{\ell(x,u)}{g(x,u)} \Gamma(\eta, \lp)
    =
    F_g(\eta) \bigg\langle \int_0^\infty \ell(x,u) du , \eta(dx) \bigg\rangle .
$$
Comparing this to the definition of $A$
(equation~\eqref{eqn:limiting_lookdown_generator}),
we see that
\begin{align*}
    \int Af(\lp) \Gamma(\eta,d\lp)
    &=
    F_g(\eta) \bigg\langle
        \int_0^\infty
        (\ell_1(x,u) + \ell_2(x,u) + \ell_3(x,u))
        du, \eta(dx)
    \bigg\rangle ,
\end{align*}
where
\begin{align*}
    \ell_1(x,u)
    &=
    \gamma(x,\eta)\left(
        \DG(g(\cdot,u) r(\cdot,\eta))(x) - g(x,u) \DG r(x,\eta)
    \right) \\
    &=
    \gamma(x,\eta)\left(
        \DG((g(\cdot,u)-1) r(\cdot,\eta))(x) - (g(x,u)-1) \DG r(x,\eta)
    \right)
\end{align*}
and
\begin{align*}
    \ell_2(x,u)
    =
    2 g(x,u) \alpha \gamma(x,\eta) r(x,\eta) \int_u^\infty(g(x,v) - 1)dv
\end{align*}
and
\begin{align*}
    \ell_3(x,u)
    =
    \left( \alpha \gamma(x,\eta)r(x,\eta)u^2 - \{\gamma(x,\eta)\DG r(x,\eta) + F(x,\eta)\} u \right)
    \partial_u g(x,u) .
\end{align*}

First note that since $\DG$ acts on space, it commutes with the integral over levels, and so
\begin{align*}
    \int_0^\infty \ell_1(x,u) du
    &=
    \gamma(x,\eta)\left(
        \DG(h(\cdot) r(\cdot,\eta))(x) - h(x) \DG r(x,\eta)
    \right) ,
\end{align*}
since $h(x) = \int_0^\infty (g(x,u) - 1) du$.
Next,
\begin{align*}
    \int_0^\infty \ell_2(x,u) du
    &=
    \alpha \gamma(x,\eta) r(x,\eta) 2\int_0^\infty g(x,u) \int_u^\infty (g(x,v)-1) dv du .
\end{align*}
Finally, integrating by parts,
\begin{align*}
    \int_0^\infty \ell_3(x,u) du
    &=
    - \alpha \gamma(x,\eta) r(x,\eta) \int_0^\infty 2u(g(x,u) - 1) du
    \\ & \qquad {}
    + \{ \gamma(x,\eta) \DG r(x,\eta) + F(x,\eta) \} h(x)
\end{align*}
Now, note that
\begin{align*}
&
    \int_0^\infty g(x,u) \int_u^\infty (g(x,v)-1) dv du - \int_0^\infty u(g(x,u)-1)du
\\ & \qquad =
    \int_0^\infty g(x,u) \int_u^\infty (g(x,v)-1) dv du - \int_0^\infty \int_v^\infty (g(x,u) -1) du dv
\\ & \qquad =
    \int_0^\infty (g(x,u)-1) \int_u^\infty (g(x,v)-1) dv du
\\ & \qquad =
    h(x)^2/2 .
\end{align*}

Adding these together, we get that
\begin{align*}
    &
    \int_0^\infty (\ell_1(x,u) + \ell_2(x,u) + \ell_3(x,u)) du
    \\ &\qquad=
    \gamma(x,\eta) \DG(h(\cdot)r(\cdot,\eta))(x)
    + F(x,\eta) h(x) 
    + \alpha \gamma(x,\eta) r(x,\eta) h(x)^2 ,
\end{align*}
which agrees with \eqref{eqn:pgen_infty}, as desired.
\end{proof}
